---
title: "Choosing StrathE2E domain extent"
author: "Jack H Laverick, University of Strathclyde"
date: "`r format(Sys.time(), '%A %d %b')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center")

Tidy_packages <- c("tidyverse", "data.table", "pbapply", "furrr", "viridis", "patchwork") # List handy data packages
Geo_packages <- c("sf", "rnaturalearth",  "raster", "stars")                # List GIS packages
lapply(c(Tidy_packages, Geo_packages), library, character.only = TRUE)      # Load packages
source("./R scripts/@_Region file.R")                                       # Define project region 

plan(multisession)                                                          # Choose the method to parallelise by

domain <- readRDS("./Objects/Domains.rds")

Bathymetry <- readRDS("./Objects/Bathymetry_points.rds") %>%                # Get bathymetry
#  .[seq(1, nrow(.), 16),] %>%                                               # Reduce resolution for plotting speed
  .[seq(1, nrow(.), 4),] %>%                                               # Reduce resolution for plotting speed
  filter(between(Elevation, -1000, 0))  

cells <- st_as_stars(Bathymetry)                                            # Convert to stars to get cells instead of points, (so no gaps) 
st_crs(cells) <- st_crs(4326)                                               # set lat-lon crs

cells <- st_as_sf(cells, as_points = F, merge = F) %>%                      # Convert the stars grid into SF polygons
  drop_na() %>% 
  st_transform(crs = crs)

world <- ne_countries(scale = "medium", returnclass = "sf") %>%             # Get a world map
  st_transform(crs = crs)                                                   # Assign polar projection


```

# Horizontal extent

We need to choose the geographic extent of our model domain. There are a number of properties which would be good to bear in mind. 

* The offshore zone should extend to the continental slope. 
* Fishing activity in the Barents Sea is high along the shelf edge, so this activity should be within in the model domain. * The deep zone is directly below the offshore zone, so limit between shallow and deep also effects the horizontal partition. 
* The depth of the shallow zone should be deep enough that there isn't significant mixing between the shallow and deep layers (see vertical boundary section).

To get a sense of some options, we plotted all the combinations of some possible minimum (30, 40, 50, 60) and maximum (300, 400, 500, 600) depths(m) for the offshore zone. 

The depth limits for the inshore zone will be 0 and the minimum depth for the offshore zone.

```{r Depth limits, eval = FALSE}

combos <- expand.grid(Shallow = c(-30,-40, -50, -60), 
                      Deep = c(-300,-400,-500, -600))                        # Create combinations of max and min depths

limit <- function(Shallow, Deep) {
  
labels <- data.frame("Deep" = abs(Deep), "Shallow" = abs(Shallow))
  
Shape <- filter(cells, between(Elevation, Deep, Shallow)) %>%
  st_union() %>%
  st_sf(labels)
  }                                      # Filter the bathymetry by a max and min depth combo and turn to a polygon

Choices <- future_map2(combos$Shallow, combos$Deep, .f=limit) %>% 
  rbindlist() %>%
  st_as_sf() %>% 
  st_transform(crs = crs)

#### Plot choice maps ####

ggplot() +
  geom_sf(data = Choices, aes(geometry = .), fill = "yellow", size = 0.01) +
  geom_sf(data = world, fill = "Black", colour = "Black", size = 0.1) +
  theme_minimal() +
  zoom +
  facet_grid(rows = vars(Shallow), cols = vars(Deep)) +
  labs(caption = "Bathymetry choice maps", x = NULL, y = NULL) +
  theme(axis.text = element_blank()) +
  NULL

    
```

We ended up choosing depth limits of 60 m and 500 m. 60 m was chosen because of vertical mixing (see the vertical mixing section). 

## Inspecting Distance to Shore

In StrathE2E terrestrial inputs can only reach the offshore zone by transiting through the inshore zone. In areas with very steep bathymetry it is possible the offshore zone polygons created above could contact land. We therefore also need to set a minimum distance from shore for the offshore zone. This ensures the inshore zone exists all along the coast. In practice this distance needs to allow a pixel of NEMO-MEDUSA model output to sit between the coast and offshore zone polygon.

```{r distance to shore, eval = FALSE}

shrink <- filter(world, subregion == "Southern Africa") # Measuring distances goes faster if we don't check every country

close <- st_as_sf(Bathymetry, coords = c("Longitude", "Latitude"), crs = 4326) %>% # Set dataframe to SF format
  st_transform(crs) 

dist <- st_distance(close, shrink) %>% pbapply(1, min)                      # Calculate the distances between points and polygons, grab the closest
close$Shore_dist <- dist                                                    # Send distances to a column

ggplot() +
  geom_sf(data = close, aes(geometry = geometry, colour = Shore_dist), size = 0.1) +
  geom_sf(data = world, size = 0.1) +
  theme_minimal() +
#  theme(axis.text = element_blank()) +
  scale_colour_viridis(name = 'Distance (m)') +
  guides(colour = guide_colourbar(barwidth = 0.5, barheight = 15)) +
  labs(caption = "Distance from shore") +
  zoom +
  NULL
```

We chose a minimum distance from shore for the offshore zone of 20km.This results in the following:

```{r domain, eval = TRUE}

colours <- c(Inshore = "yellow", Offshore = "yellow3")

ggplot() + 
  geom_sf(data = domain, aes(fill = Shore), colour = NA) +
#  geom_sf(data = Region_mask, colour = "red", fill = NA) + 
  geom_sf(data = world, size = 0.1, fill = "black") +
  scale_fill_manual(values = colours, name = "Zone") +
  zoom +
  theme_minimal() +
  labs(caption = "Suggested model area") +
  NULL

GFW <- readRDS("./Notes/Cache/domain_GFW.rds") %>% 
  st_transform(crs = crs)

ggplot() +
  geom_sf(data = world, size = 0.1, fill = "grey30", colour = "white", size = 0.1) +
  geom_sf(data = domain, fill = "grey", colour = "black", size = 0.1) +
  geom_stars(data = GFW, colour = NA) +
  geom_sf(data = domain, fill = NA, colour = "white", size = 0.2) +
  viridis::scale_fill_viridis(na.value = NA, trans = "log10", option = "inferno") +
  zoom +
  theme_minimal() +
  labs(caption = "Total fishing effort") +
  NULL

# map + fish

```

Any suggestions of where to "cut" the domain to stop it extending too far would be appreciated.

# Vertical boundary

```{r setupV, include=FALSE, eval = FALSE}

#### Get the subset of NEMO-MEDUSA pixels in the model domain ####

file <- list.files("../../../../../import/fish/South_Atlantic/", 
                   recursive = T, full.names = TRUE, pattern = "grid_W")[1]

NM_space <- raster(file, varname = "nav_lat") %>% 
  as.data.frame(xy = T) %>% 
  cbind(Lon = as.data.frame(raster(file, varname = "nav_lon"))) %>% 
  setNames(c("x", "y", "latitude", "longitude")) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_join(st_transform(domain, 4326)) %>% 
  drop_na() %>% 
  st_drop_geometry() %>% 
  dplyr::select(x,y)

setDT(NM_space, key = c("x", "y"))
```

StrathE2E contains 3 ocean volumes. There's an inshore-offshore divide, and a shallow-deep divide. Only the offshore volume is split into shallow and deep. Choosing a shallow-deep boundary depth has knock-on consequences for the inshore-offshore divide, as the inshore zone will be defined as the sea surface to this depth. StrathE2E assumes that the volumes are well mixed, and any exchange between the deep and shallow compartments are controlled by vertical diffusivity at that depth. The boundary depth therefore needs to be deep enough to avoid excessive mixing between the volumes, while also considering how large the inshore zone will end up being. Look at the "Choose our horizontal boundaries" document for more details.

We noticed the NEMO-MEDUSA model outputs contained a cluster of implausibly high values (~10). This was verified by comparing to the distribution of values from SINMOD model outputs also from the Barents Sea. Values this high aren't usable in StrathE2E so we need to come up with an approach to work around them. We could either "avoid" them by choosing an appropriate boundary depth, or we can accommodate it by choosing an overwrite value.

``` {r plot comparison cached, eval = FALSE}

a <- readRDS("./Notes/Cache/domain_V_a.rds")

a
```

## Deep convection

After asking at NOC it looks like these values are used to represent deep covection in the model. When temperatures get very cold the water column becomes unstable and just mixes right the way through from the surface to the seafloor. 0.14 was the maximum value present in SINMOD so for now we use 0.14 as the cutoff threshold to define values representing deep convection. This allows us to see how much of the model output is affected.


Below I'm showing the depth profile for an example year (1980) and how deep convection changes through the year. The white line is our currently proposed depth boundary of 60 m. The horizontal area being assessed is shown in grey (this shrinks with depth as the seabed encroaches on the ocean volume. For small areas the proportions can become sensitive).

``` {r Deep convection cached, eval = FALSE}

b <- readRDS("./Notes/Cache/domain_V_b.rds")

b
```

As expected deep convection is most noticeable in the shallows and in January:March (Q1) as opposed to July:September (Q3). 60m is free of deep convection in Q3, but for other times of year it looks impossible to avoid the process.

## Vertical diffusivity 

Ignoring deep convection, it is still interesting to see if there are any patterns in vertical diffusivity with depth and time of year. A plot similar to the one above is shown below for mean vertical diffusivity (ignoring deep convection). Because deep convection has been removed you'll notice a different shape to the grey area in the background. We now have the largest amount of data at ~130 m as at depth the seabed encroaches, and in the shallows deep convection is common. 

``` {r Vertical diffusivity cached, eval = FALSE}

c <- readRDS("./Notes/Cache/domain_V_c.rds")

c
```

Again water is more mixed in winter than in summer. 60m looks like a pretty good boundary as values level off afterwards for all times of year.

